Stochastic Approximation Monte Carlo with a Dynamic Update Factor

### 2019-07-23 papers/histogram/sad-referee-response.md

...

Response to first referee

The first referee acknowledges that our paper is well written and that the field
is an important one, but does not believe that our work is a substantial
contribution to the field.  The referee has three major criticisms of the value
of our work (aside from issues with the introduction section).

***Jordan Says:*** Computational Physics Papers in Physical Review E
Appropriate papers for this section are those that report new advances in
computational physics. This includes papers that describe new developments of
computational methods or substantial improvements of existing computational
methods, provided they are applicable to significant physics problems.
In many cases such a paper will improve its impact by explicitly demonstrating
the usefulness of the described method, i.e., by applying it to some physics
problem (old or new). However, it should be noted that such an application is
not a necessary condition for eligibility. Rather, the requirement that "papers
must contain new results in physics" should be interpreted as a requirement of
novelty with respect to the method(s), the physics, or both.
*PRE only even loosely requires that new methods be tested on a single (old) or new system.
Simply arguing that our method is distinct from WL (the referees argument) 'it
is a new version of SAMC!' should be sufficient based on PRE guidelines.*

@@

Firstly, the referee criticizes our paper for testing the method on a single
physical system (the square well fluid).  The referee gives as an argument an
example of a paper which tests a method on three different physical systems.  We
give as counter examples references **look them up** (including references
**look them up** which the referee asked us to add) which each use a single
physical system as the test case.  We agree that our method should be tested on
additional different physical systems but dispute that this is necessary for the
first paper introducing this method.  **Decide whether to just throw in the
Ising...**

**look them up**
It is quite common for papers detailing comparison of convergence for various
methods to only examine a single physical or sometimes even an unphysical (as in
the case of integrals) system. When examining the convergence and/or comparing
various computational methods it cannot therefore be expected that multiple
systems should be considered in the same work. The authors provide a list of
such works that they directly cite (many accepted by PRE) that prove their point:

Analysis of the convergence of the 1/t and Wang-Landau algorithms in the calculation
of multidimensional integrals (accepted by PRE)
This paper compares WL and 1/t-WL on several different integrals (not even a real physical system because the exact answer is known)

Avoiding boundary effects in Wang-Landau sampling (accepted by PRE)
This paper with DP Landau as an author only considers the 2D ising model with
the novelty of the paper to avoid boundary effects in WL which is succinctly
stated in the abstract: 'A simple modification of the ‘‘Wang-Landau sampling’’
algorithm...'

Optimal modification factor and convergence of the Wang-Landau algorithm (accepted by PRE)
This paper examines the optimal modification factor and tests only on the 2D
ising model.

Convergence of Stochastic Approximation Monte Carlo and modified
Wang–Landau algorithms: Tests for the Ising model (accepted by Computer Physics Communications)
This paper compares SAMC, WL, and 1/t-WL on a single system: the 2D ising model

Stochastic approximation Monte Carlo and Wang–Landau Monte Carlo applied to a continuum polymer model (accepted by Computer Physics Communications)
SAMC and WL are compared for a single continuum model with a square well type potential.

The referee even suggested to cite some of the original multicannonical literature
Multicanonical Ensemble: A New Approach to Simulate First-Order Phase Transitions (accepted by PRL)
This important work considers a single system in the 2D 10 state Potts model.

@@

Thirdly, because we find that 1/t-WL converges at the same rate as our method,
the referee concludes that our method is not a substantial improvement. We note
first (as mentioned in the paper in **...**) that the 1/t-WL method is solving a
significantly easier problem, because it is told in advance the range of
energies that are of interest.  We agree, however, that the benefit of our
method lies specifically in the fact that a temperature range is given as an
input.  Achieving the same result as the SAD method with 1/t-WL would either
require multiple simulations over a wider range of energies, or would require
guessing in advance a wider range of energies than are needed.  In either case,
the convergence would be worse than our tests indicate.

1. **Existence of the SAMC method.** We state in the paper "In this work, we
have developed an improved algorithm based on SAMC."  This is again reinforced
in Secion III.

2. We have added a section detailing the important impact of multicanonical
methods on influencing methods like WL etc... referencing the original authors
to the introduction.

3. As per the referees request, we have briefly reviewed methods the Statistical
temperature Monte Carlo method.

4.  As per the referees request, we have briefly reviewed methods used in the
metadynamics communities in the introduction.

5. more test cases

@@

Minor details:

6. Out of curiosity: The authors say a temperature range T_min < T <
infinity should be chosen. Why not T_min < T < T_max?

7. On page 5 the authors say "SAD should perform similarly ...". Is
there any evidence?

### 2019-07-25 papers/histogram/sad-referee-response.md

...

Response to first referee

The first referee acknowledges that our paper is well written and that the field
is an important one, but does not believe that our work is a substantial
contribution to the field.  The referee has three major criticisms of the value
of our work (aside from issues with the introduction section).

***Jordan Says:*** Computational Physics Papers in Physical Review E
Appropriate papers for this section are those that report new advances in
computational physics. This includes papers that describe new developments of
computational methods or substantial improvements of existing computational
methods, provided they are applicable to significant physics problems.
In many cases such a paper will improve its impact by explicitly demonstrating
the usefulness of the described method, i.e., by applying it to some physics
problem (old or new). _However, it should be noted that such an application is
not a necessary condition for eligibility_. Rather, the requirement that "papers
must contain new results in physics" should be interpreted as a requirement of
novelty with respect to the method(s), the physics, or both.
*PRE only even loosely requires that new methods be tested on a single (old) or new system.
Simply arguing that our method is distinct from WL (the referees argument) 'it
is a new version of SAMC!' should be sufficient based on PRE guidelines.*

Firstly, the referee criticizes our paper for testing the method on a single
physical system (the square well fluid).  The referee gives as an argument an
example of a paper which tests a method on three different physical systems.  We
give as counter examples references **look them up**  which each use a single
physical system as the test case.  We agree that our method should be tested on
additional different physical systems but dispute that this is necessary for the
first paper introducing this method.  **Decide whether to just throw in the
Ising...**
It is quite common for papers detailing comparison of convergence for various
methods to only examine a single physical or sometimes even an unphysical (as in
the case of integrals) system. When examining the convergence and/or comparing
various computational methods it cannot therefore be expected that multiple
systems should be considered in the same work. The authors provide a list of
such works that they directly cite (many accepted by PRE) that prove their point:

Analysis of the convergence of the 1/t and Wang-Landau algorithms in the calculation
of multidimensional integrals *(accepted by PRE)*
This paper compares WL and 1/t-WL on several different integrals (not even a
real physical system because the exact answer is known)

Avoiding boundary effects in Wang-Landau sampling *(accepted by PRE)*
This paper with DP Landau as an author only considers the 2D ising model with
the novelty of the paper to avoid boundary effects in WL which is succinctly
stated in the abstract: 'A simple modification of the ‘‘Wang-Landau sampling’’
algorithm...'

Optimal modification factor and convergence of the Wang-Landau algorithm *(accepted by PRE)*
This paper examines the optimal modification factor and tests on a single
system: the 2D ising model.

Convergence of Stochastic Approximation Monte Carlo and modified
Wang–Landau algorithms: Tests for the Ising model *(accepted by Computer Physics Communications)*
This paper compares SAMC, WL, and 1/t-WL on a single system: the 2D ising model

Stochastic approximation Monte Carlo and Wang–Landau Monte Carlo applied to a continuum polymer model *(accepted by Computer Physics Communications)*
SAMC and WL are compared for a single continuum model with a square well type potential.

@@

Secondly, the referee does not see much value in providing a temperature range
(rather than an energy range) as input to the method.  We believe that a
temperature range is *far* more physical and convenient, and this is the primary
value of our paper.  **We have modified the abstract to better highlight this
contribution.**

Thirdly, because we find that 1/t-WL converges at the same rate as our method,
the referee concludes that our method is not a substantial improvement. We note
first (as mentioned in the paper in **...** The results section we could highlight again in the conclusion?) that the 1/t-WL method is solving a
significantly easier problem, because it is told in advance the range of
energies that are of interest.  We agree, however, that the benefit of our
method lies specifically in the fact that a temperature range is given as an
input.  Achieving the same result as the SAD method with 1/t-WL would either
require multiple simulations over a wider range of energies, or would require
guessing in advance a wider range of energies than are needed.  In either case,
the convergence would be worse than our tests indicate.

@@

1. **Existence of the SAMC method.** We state in the paper "In this work, we
have developed an improved algorithm based on SAMC."  This is again reinforced
in Secion III.

Multidimensional stochastic approximation Monte Carlo
*(accepted by PRE)*
"the Wang-Landau MC [3] scheme and its mathematical generalization, stochastic approximation
MC (SAMC) [4–6].""

Convergence estimation of flat-histogram algorithms based on simulation results
*(accepted by Computer Physics Communications)*
"At the same time, Liang suggested a generalized simulation method:stochastic
approximation Monte Carlo (SAMC) [7,8] including the 1/t-WL as a particular case"

Stochastic approximation Monte Carlo and Wang–Landau Monte Carlo applied to a continuum polymer model
*(accepted by Computer Physics Communications)*
"Liang et al.[30] showed that WLMC could be seen as a version of SAMC and using
the mathematical background of stochastic approximation methods they proved the
convergence of SAMC."

### 2019-08-01 papers/histogram/sad.tex
various paper(s) edits.

\section{Introduction}
Over the past several decades a number of flat histogram Monte Carlo simulation
algorithms have been developed which calculate the thermodynamic properties of
various systems for over a range of temperatures.  This development began with
the original histogram method, which used a single canonical Monte Carlo
simulation to predict properties for nearby
temperatures~\cite{ferrenberg1988new}.  For large systems this approach is
limited to a narrow temperature range because a single canonical simulation
explores only a small range of energies.  This led to a variety of ``flat'' (or
``broad'') histogram methods~\cite{penna1996broad, penna1998broad,
swendsen1999transition, wang2001determining, wang2001efficient,
trebst2004optimizing}, which explore a wider range of energies.  Another
benefit, in contrast with low-temperature canonical Monte Carlo, is that these
approaches cannot be trapped in a local energy minimum.

Wang and Landau introduced the most widely used flat histogram algorithm (WL)
that uses an update factor and a statistical histogram to compute the density of
states of a given system~\cite{wang2001determining, wang2001efficient}.  While
the method is incredibly powerful, it has a few disadvantages. Firstly, it
requires the user to select the range of energies to be
studied~\cite{landau2004new}. This adds an additional hurdle to its application
to systems for which the interesting range of energies is not known \emph{a
priori}.  The simulation violates detailed balance, albeit briefly as the size of
the violation decreases with time, which complicates convergence analysis.  In
fact, the error in a WL computation has been demonstrated to saturate
at a non-zero value~\cite{yan2003fast}, i.e. the method does not converge to the
true density of states~\cite{belardinelli2007wang, belardinelli2007fast,
belardinelli2008analysis, belardinelli2014intrinsic, singh2012density}.
%
Belardinelli and Pereyra demonstrated that allowing the update factor to
decrease faster than $1/t$ leads to nonconvergence~\cite{belardinelli2007wang}
where $t$ corresponds to the number of moves.
%
This leads to their $1/t$-WL algorithm which ensures that the error continues to
decrease asymptotically as $1/\sqrt{t}$~\cite{belardinelli2008analysis}.

@@

In this work, we have developed an improved algorithm based on SAMC that does
not require an array of non-physical, user-defined inputs and therefore should
be easily applicable to any system. The method does require the user to define a
temperature range of interest ($T_\text{min}$ to $T_\infty$).  We call this
method SAD (Stochastic Approximation with a Dynamic update factor), and will
discuss it in detail in the methods section. We compare its convergence
properties with three existing flat histogram methods: WL, $1/t$-WL, and SAMC.

@@

\section{Flat histogram methods}\label{sec:histogram}

The goal of flat histogram methods (also called \emph{broad histogram}
or \emph{multicanonical} methods) is to simulate each energy with
similar accuracy, so as to accurately determine the density of states
over a broad range of energies---and thus to determine the thermodynamic
quantities such as heat capacity or internal energy
over a broad range of temperatures.
Properties that require more information---such as a spatial
correlation function or a response function---can still be computed
for any temperature, provided statistics are collected for each
individual energy, which can then be reweighted for any
temperature~\cite{errington2003direct}.

@@

Flat histogram methods calculate the density of states $D(E)$ for a
discrete set of energies~\cite{haber2018performance}. For this reason,
energy binning becomes an important consideration for systems with a
continuum of possible energies.  Energy bins are typically of uniform
size for the entire energy continuum~\cite{fasnacht2004adaptive}. Some
methods such as AdaWL~\cite{koh2013dynamically} employ a tunable
mechanism for controlling the binning for low entropic states in order
to ensure the exploration of all energies.  The method introduced in
this paper is designed to scale appropriately as bin size is changed,
but we do not test this scaling, as we use a system with discrete
energy levels.

In this section we will introduce four closely related flat histogram
methods which each rely on a weight function $w(E)$.  In these
algorithms, the probability of accepting a move is given by
\begin{equation}
	\mathcal{P}(E_\text{old} \rightarrow E_\text{new})
	= \min\left[1,\frac{w(E_\text{old})}{w(E_\text{new})}\right]
\end{equation}
which biases the simulation in favor of energies with low weights.  A
set of weights that are proportional to the density of states of the
system $D(E)$ will result in an entirely flat histogram.  Thus flat
histogram is a criteria for convergence for these methods.  To avoid
overflow error, since the weights may vary over more than a few
hundred orders of magnitude, the natural logarithm of the weights are
stored.  Since in the microcanonical ensemble the entropy is defined
as $S(E) \equiv \ln(D(E))$ (where $k_B \equiv 1$), the logarithm of the weights can be
thought of as an approximation of the entropy.

Each approach uses a random walk in energy space to
estimate the density of states.  The core of these approaches
is to continuously update the weights at each step of the simulation
\begin{equation}
	\ln{w_{t+1}(E)}=\ln{w_{t}(E)}
	+\gamma_t
\end{equation}
where $t$ is number of the current move, $\gamma_t(t)$ is a
move-dependent update factor, and $E$ is the current energy.  This
update causes the random walk to avoid energies that have been
frequently sampled, leading to a rapid exploration of energy space.
This approach, however, violates detailed balance, due to the
acceptance probabilities changing with each move.  The severity of
this violation decreases as we decrease $\gamma_t$.  The four methods
differ primarily in how they schedule the decreasing of $\gamma_t$.

@@

The Wang-Landau approach~\cite{wang2001efficient,wang2001determining,
  landau2014guide} begins with $\gamma_t=1$, and then decreases $\gamma$
in discrete stages.  We track the number of moves ending at each
energy in a histogram.  When that histogram is sufficiently flat,
$\gamma_t$ is decreased by a specified factor of $\frac12$.  The
flatness is defined by the ratio between the minimum value of the
histogram and its average value.  When this flatness reaches a
specified threshold (typically 0.8), the $\gamma_t$ value is decreased
and the histogram is reset to zero.  This approach requires that the
energy range of interest be known in advance, and difficulties can
occur with this flatness criteria due to the fact that some energies
in this energy range might never be
sampled~\cite{haber2014transition}.  The entire process is repeated
until $\gamma_t$ reaches a desired cutoff.

@@

While this approach is very efficient and has been widely used, it
suffers a few shortcomings.  Firstly, the set of energies must be
specified ~\cite{wang2001efficient, schulz2003avoiding, yan2003fast},
which may require multiple simulations. Secondly, while Wang-Landau
converges quickly, it does not in general converge to the true density
of states ~\cite{belardinelli2008analysis, zhou2008optimal}.  It can
and does decrease $\gamma_t$ so quickly that it will never (for any
cutoff value) decrease the error in the density of states beyond a
given nonzero value.

\subsection{$1/t$-Wang-Landau}
The error saturation of Wang-Landau algorithm algorithm can be
corrected by modifying the update factor such that it does not
decrease too quickly.  Belardinelli and Pereyra raised the question as
to whether the update factor should be decrease by $\nicefrac12$ or
some other factor for each update~\cite{belardinelli2007fast}. They
implemented a schedule that enforced that if the update factor
$\gamma_t$ is ever less than $\nicefrac{1}{t}$ then the update factor
is set to $\nicefrac{1}{t}$ and the histogram is no longer tracked. Employing this
scheduler, they found that the error saturation is avoided since the
correct density of states is approached asymptotically as
$t^{-\frac12}$~\cite{belardinelli2008analysis}. Zhou et. al further
confirmed that the WL algorithm never converges exponentially and
succesfully bounded the statistical error between $t^{-\frac12}$ and
$1/t$~\cite{zhou2008optimal}.

Schneider et. al outline minor refinements to the $1/t$-WL algorithm
including $\nicefrac{N_S}{t}$ scaling and switching from standard WL
to $1/t$-WL when the update factor $\gamma_t <
\nicefrac{N_S}{t}$~\cite{schneider2017convergence}. As per the
original $1/t$ implementation~\cite{belardinelli2007fast}, the update
factor is decreased once all $N_S$ energy states are visited at least
once (i.e. $H(E) \neq 0$) effectively avoiding the concept of
`flatness'.  In this work, we refer to this refined algorithm as $1/t$-WL,
with the update factor asymptoting to $\nicefrac{N_S}{t}$.

@@

One might reasonably choose to err by selecting a large $t_0$.
The rate of convergence is harder to estimate when $t_0$ is large, but
in general $\gamma_t$ itself forms a lower bound on the accuracy with which
the entropy may be known, with an unknown prefactor which is roughly
the coherence time of the Monte Carlo simulation.  Since $\gamma_t$ is
given by $t_0/t$, the time to converge to a given accuracy is increased
in proportion to the ratio by which we overestimate $t_0$.  Thus, while
it is exponentially painful to underestimate $t_0$, overestimating by
several orders of magnitude is also not acceptable.  We should note
that these extreme limiting cases do not preclude the possibility that there
is a wide range of $t_0$ values that lead to an acceptable convergence
rate.

@@

@@

\section{Conclusions}
We have introduced a new algorithm that effectively samples the
energy space corresponding to a desired range of temperatures for a
variety of system sizes.
%
We find that both SAD and $1/t$-WL demonstrate excellent and robust
convergence.
They both converge more rapidly than SAMC, and unlike WL consistently
converge to the correct density of states. SAD requires
the user to specify a temperature range of interest rather than an
energy range of interest as $1/t$-WL does.  For use cases in which
a range of desired temperatures is known, this will make the SAD method
considerably more convenient.

>>>>

\section{Introduction}
Over the past several decades a number of flat histogram Monte Carlo simulation
algorithms have been developed which calculate the thermodynamic properties of
various systems for over a range of temperatures.  This development began with
the original histogram method, which used a single canonical Monte Carlo
simulation to predict properties for nearby
temperatures~\cite{ferrenberg1988new}.  For large systems this approach is
limited to a narrow temperature range because a single canonical simulation
explores only a small range of energies. Berg and Neuhaus developed multicanonical
methods that introduced a weight function to enable flat histogram sampling
which improved the exploration of configuration space and allows the system
to
overcome free energy barriers~\cite{berg1991multicanonical, berg1992multicanonical}.
These works led to an explosion in the interest of
``flat'' (or
``broad'') histogram methods~\cite{penna1996broad, penna1998broad,
swendsen1999transition, wang2001determining, wang2001efficient,
trebst2004optimizing}, which explore a wider range of energies.  Another
benefit, in contrast with low-temperature canonical Monte Carlo, is that these
approaches cannot be trapped in a local energy minimum.

Wang and Landau introduced the most widely used flat histogram algorithm (WL)
that uses an update factor and a statistical histogram to compute the density of
states of a given system~\cite{wang2001determining, wang2001efficient}.  While
the method is incredibly powerful, it has a few disadvantages. Firstly, it
requires the user to select the range of energies to be
studied~\cite{landau2004new}. This adds an additional hurdle to its application
to systems for which the interesting range of energies is not known \emph{a
priori}.  The simulation violates detailed balance, albeit briefly as the size of
the violation decreases with time, which complicates convergence analysis.  In
fact, the error in a WL computation has been demonstrated to saturate
at a non-zero value~\cite{yan2003fast}, i.e. the method does not converge to the
true density of states~\cite{belardinelli2007wang, belardinelli2007fast,
belardinelli2008analysis, belardinelli2014intrinsic, singh2012density}.
%

Belardinelli and Pereyra demonstrated that allowing the update factor to
decrease faster than $1/t$ leads to nonconvergence~\cite{belardinelli2007wang}
where $t$ corresponds to the number of moves.
%
This leads to their $1/t$-WL algorithm which ensures that the error continues to
decrease asymptotically as $1/\sqrt{t}$, which they demonstrated avoids the
error saturation and asymptotically approaches the true density of
states~\cite{belardinelli2008analysis}.
Zhou et. al further
confirmed that the original WL algorithm never converges exponentially and
succesfully bounded the statistical error between $t^{-\frac12}$ and
$1/t$~\cite{zhou2008optimal}.
Schneider \emph{et al.} outline minor refinements to the $1/t$-WL algorithm
including scaling the update factor with the number of energy bins~\cite{schneider2017convergence}.

@@

Kim \emph{et al.} introduced Statistical Temperature Monte Carlo (STMC) and
the related Statistical Temperature Molecular Dynamics (STMD) which is an
improvement on the WL method which approximates the entropy (or log density of states)
as a piecewise
linear function, which improves convergence for systems with a continuously varying
energy~\cite{kim2006statistical, kim2007statistical}. STMC
applied to WL requires a temperature range be specified rather than an energy
range.  Kim \emph{et al.} extended this work as Replica Exchange Statistical
Temperature Monte Carlo (RESTMC), which uses replica exchange of multiple overlapping
STMC simulations to improve convergence~\cite{kim2009replica}.
More recently, Junghans \emph{et al.} have demonstrated a close
connection between metadynamics, which was introduced by Laio and
Parinello~\cite{laio2002escaping}, and WL-based Monte Carlo
methods, with STMD forging the connection~\cite{junghans2014molecular}.

In this work, we have developed an improved algorithm based on SAMC that does
not require an array of non-physical, user-defined inputs and therefore should
be easily applicable to any system. The method (like STMC above) does require the user to define a
temperature range of interest ($T_\text{min}$ to $T_\infty$).  We call this
method SAD (Stochastic Approximation with a Dynamic update factor), and will
discuss it in detail in the methods section. We compare its convergence
properties with three existing flat histogram methods: WL, $1/t$-WL, and SAMC.

@@

\section{Flat histogram methods}\label{sec:histogram}
\jpsays{REFEREE 2: The description of flat histogram methods in section II was a bit hard
to understand at first and may need to be edited to better highlight
the differences between the methods. Also, the entropy argument in
section II.D is unclear. The relationship between entropy, the
histogram, and the update factor needs to be more fully described and
the basis for equation 7 (and the approximations made there) needs to
be made clearer.}

The goal of flat histogram methods (also called \emph{broad histogram}
or \emph{multicanonical} methods) is to simulate each energy with
similar accuracy, so as to accurately determine the density of states
over a broad range of energies---and thus to determine the thermodynamic
quantities such as heat capacity or internal energy
over a broad range of temperatures.
Properties that require more information---such as a spatial
correlation function or a response function---can still be computed
for any temperature, provided statistics are collected for each
individual energy, which can then be reweighted for any
temperature~\cite{panagiotopoulos1998phase, panagiotopoulos2000monte,
errington2003direct}.

@@

Flat histogram methods calculate the density of states $D(E)$ for a discrete set
of energies~\cite{wang2001determining, dayal2004performance, troyer2003flat,
trebst2004optimizing}. For this reason, energy binning becomes an important
consideration for systems with a continuum of possible energies.  Energy bins
are typically of uniform size for the entire energy
continuum~\cite{fasnacht2004adaptive}. Some methods such as
AdaWL~\cite{koh2013dynamically} employ a tunable mechanism for controlling the
binning for low entropic states in order to ensure the exploration of all
energies.  The method introduced in this paper is designed to scale
appropriately as bin size is changed, but we do not test this scaling, as we use
a system with discrete energy levels.

In this section we will introduce four closely related flat histogram
methods each of which rely on a weight function $w(E)$.  In these
algorithms, the probability of accepting a move is given by
\begin{equation}
	\mathcal{P}(E_\text{old} \rightarrow E_\text{new})
	= \min\left[1,\frac{w(E_\text{old})}{w(E_\text{new})}\right]
\end{equation}
which biases the simulation in favor of energies with low weights.  A
set of weights that are proportional to the density of states of the
system $D(E)$ will result in an entirely flat histogram, which means that the
weights should converge to being proportional to the density of states.  The
natural logarithm of the weights are typically stored, since the density of
states will often vary over a few hundred orders of magnitude.
In the microcanonical ensemble the entropy is defined
as $S(E) \equiv \ln(D(E))$ (where $k_B \equiv 1$), the logarithm of the weights
is an approximation of the entropy.

Each approach uses a random walk in energy space to
estimate the density of states.  The core of these approaches
is to update the weights at each step of the simulation
\begin{equation}
	\ln{w_{t+1}(E)}=\ln{w_{t}(E)}
	+\gamma_t
\end{equation}
where $t$ is number of the current move, $\gamma_t(t)$ is a
update factor that varies over the simulation, and $E$ is the current energy.  This
update causes the random walk to avoid energies that have been
frequently sampled, enabling a rapid exploration of energy space.
This approach violates detailed balance, due to the
acceptance probabilities changing with each move, but the severity of
this violation decreases as we decrease $\gamma_t$.  The four methods
differ primarily in how they schedule the decrease of $\gamma_t$.

@@

The Wang-Landau approach~\cite{wang2001efficient,wang2001determining,
  landau2014guide} begins with $\gamma_t=1$, and then decreases $\gamma$
in discrete stages.  When the energy histogram is sufficiently flat,
$\gamma_t$ is decreased by a specified factor of $\frac12$.  The
flatness is defined by the ratio between the minimum value of the
histogram and its average value.  When this flatness reaches a
specified threshold (typically 0.8), the $\gamma_t$ value is decreased
and the histogram is reset to zero.  This approach requires that the
energy range of interest be known in advance, and difficulties can
occur with this flatness criteria due to the fact that some energies
in this energy range might never be
sampled~\cite{troster2005wang}.  The entire process is repeated
until $\gamma_t$ reaches a desired cutoff.

@@

\jpsays{remove and add to intro as this does not convey how to implement WL!}
While this approach is very efficient and has been widely used, it
suffers a few shortcomings.  Firstly, the set of energies must be
specified ~\cite{wang2001efficient, schulz2003avoiding, yan2003fast},
which may require multiple simulations. Secondly, while Wang-Landau
converges quickly, it does not in general converge to the true density
of states ~\cite{belardinelli2008analysis, zhou2008optimal}.  It can
and does decrease $\gamma_t$ so quickly that it will never (for any
cutoff value) decrease the error in the density of states beyond a
given nonzero value.

\subsection{$1/t$-Wang-Landau}

The $1/t$-WL algorithm ensures convergence by preventing the $\gamma_t$ factor
from dropping below $N_S/t$~\cite{belardinelli2008analysis, schneider2017convergence}.
The method follows the standard WL algorithm with two modifications.  Firstly, the histogram
is considered flat, and $\gamma_t$ is decreased by a factor of two, when every
energy state has been visited once, i.e. when the WL ``flatness'' becomes nonzero.
Secondly, when the $\gamma_t$ becomes less than $N_S/t$, the algorithm switches
to use
\begin{align}
  \gamma_t = \begin{cases}
     \gamma_{WL} & \gamma_{WL} > \frac{N_S}{t} \\
     \frac{N_S}{t} & \text{forever more}
 \end{cases}
\end{align}
where $t$ is the number of moves and $N_S$ is the number of energy bins.
\jpsays{David and I got to here!}

@@

One might reasonably choose to err by selecting a large $t_0$.
The rate of convergence is harder to estimate when $t_0$ is large, but
in general $\gamma_t$ itself forms a lower bound on the accuracy with which
the entropy may be known, with an unknown prefactor which is roughly
the coherence time of the Monte Carlo simulation \jpsays{Michael Perlin asked about this in his comments and my understanding was that we meant the time for $\gamma_t$ to reach 1/t behavior. Since this is part of a section needing editing I thought we should clarify}.  Since $\gamma_t$ is
given by $t_0/t$, the time to converge to a given accuracy is increased
in proportion to the ratio by which we overestimate $t_0$.  Thus, while
it is exponentially painful to underestimate $t_0$, overestimating by
several orders of magnitude is also not acceptable.  We should note
that these extreme limiting cases do not preclude the possibility that there
is a wide range of $t_0$ values that lead to an acceptable convergence
rate.

@@

We set the maximum temperature of interest as $T=\infty$ for several reasons.
Firstly, when this temperature is reached, the system is maximally randomized.
Allowing infinite temperature further ensures that th system can surmount any
size of energy barrier.  Finally, the high temperatures are ``easy'' to simulate,
and tend to converge very quickly.  We note, however, that this is under the
assumption that infinite temperature corresponds to a finite energy, as is the
case for configurational sampling.  Were we to include kinetic energy in a
simulation (as in molecular dynamics), or to examine a system such as a high
density Lennard-Jones fluid which might have an energy of maximum entropy far
from the energy range of physical interest, then we would need to apply a maximum
temperature, which could slow down the rate at which state space is explored.

@@

\section{Conclusions}
\jpsays{We have introduced SAD, a new variant of SAMC, that}
We have introduced a new algorithm that effectively samples the
energy space corresponding to a desired range of temperatures for a
variety of system sizes.
%
We find that both SAD and $1/t$-WL demonstrate excellent and robust
convergence.
They both converge more rapidly than SAMC, and unlike WL consistently
converge to the correct density of states. SAD requires
the user to specify a temperature range of interest rather than an
energy range of interest as $1/t$-WL does.  For use cases in which
a range of desired temperatures is known, this will make the SAD method
considerably more convenient.
